\Large{Research Seminar Software Science} \\ \huge{Simulation of monadic effects in Haskell}

### 2018-04-16 paper.tex
modified introduction, changed title of paper

\title{\Large{Research Seminar Software Science} \\
\huge{Monads}}
\author{Brigel Pineti}
\date{April 2018}

@@

@@

The code snippets presented in this paper are implemented with Haskell.

@@

In this section are reflected two main features concerning the above mentioned functional language. Taking into consideration that not every reader might be familiar with the language, I found it necessary to briefly illustrate the laziness and purity of Haskell.

@@

Haskell is a functional language with non-strict semantics. What does this mean? An expression is evaluated from the outside in. For instance, the evaluation of \textit{$(a + (b * c))$} will firstly reduce the \textit{$+$} operator and then the inner bracket \textit{$(a * b)$}. \\
Lazy evaluation is just an implementation approach for a non-strict language. An appropriate definition for laziness can be represented by the term \textit{call-by-need}. Consequently, the evaluation of an expression is delayed until its result is actually needed from another computation. Therefore, the parameters of a function are evaluated only when they are needed and not a moment before.

>>>>

\title{\Large{Research Seminar Software Science} \\
\huge{Simulation of monadic effects in Haskell}}
\author{Brigel Pineti}
\date{April 2018}

@@

\section*{Abstract}

@@

Pure function languages like Haskell rule out side effects. Why? Firstly, to allow unrestricted application of program transformation and equation logic. Secondly, being a non-strict language, Haskell's order of side effects is undefined. The aim of this paper is to show how monads can be used to exploit impure effects into a pure language. The code snippets attached to this paper are implemented in Haskell. Section 2 will provide an overview on the characteristics of the language being used. Section 3 is using the Maybe Monad as an illustrative example to get the reader's feet wet with the complex concept of a Monadic value. The application of monads is demonstrated with three case studies.

@@

In this section are reflected two main features concerning the above mentioned functional language. Taking into consideration that not every reader might be familiar with the properties of the language, I found it necessary to briefly illustrate the laziness and purity of Haskell.

@@

Haskell is a functional language with non-strict semantics. What does this mean? An expression is evaluated from the outside in. For instance, the evaluation of \textit{$(a + (b * c))$} will firstly reduce the \textit{$+$} operator and then the inner bracket \textit{$(a * b)$}. \\
Lazy evaluation is just an implementation approach for a non-strict language. An appropriate definition for laziness can be represented by the term \textit{call-by-need}. Consequently, the evaluation of an expression is delayed until its result is actually needed from another computation. Therefore, the parameters of a function are evaluated only when they are needed and not a moment before. An advantage of laziness
